Run some checks

### 2019-04-25 scripts/model/1_checks.Rmd
Added slides for NCEAS

---
title: "Untitled"
author: "Villaseñor-Derbez J.C."
date: "22/4/2019"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---
# Setup

@@

```{r}
# Load libraries
library(startR)
library(here)
library(cowplot)
library(tidyverse)

# Load functions
source(here("scripts", "model", "grow.R"))
source(here("scripts", "model", "E_vec_fxn.R"))
source(here("scripts", "model", "pi_of_E.R"))
source(here("scripts", "model", "max_Ei.R"))
source(here("scripts", "model", "solve_pi.R"))
```

@@

```{r}
left_join(act, catches, by = c("year", "eez_iso3")) %>%
  drop_na() %>% 
  ggplot(mapping = aes(x = days, y = catches, color = year)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 4) +
  labs(x = "Effort (days)", y = "Catches (tonnes)") +
  ggtitle("Relationship between effort and catches")
```

```{r}
fvec_cpue <- c(cpue$cpue_prop, prop_hs)
names(fvec_cpue) <- c(levels(cpue$eez_iso3), "HS")

@@

# Run some checks

```{r}
# Is there a value of E for the given possible ranges of Biomass?
tibble(X = seq(1, Bnow, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = sum(max_Ei(p = p, q = q, Xvec = X, theta = theta, R = R, beta = beta, c = c))) %>%
  ggplot(aes(x = X, y = E)) +
  geom_line() +
  geom_vline(xintercept = Bmsy) +
  geom_hline(yintercept = E)
```

```{r}
# What's the aggregate demand for a given value of pi with Bmsy?
# FIrst we need to establish the maximum chargable price to plot all ranges from 0 to max_pi
max_pi <- max((p * q * Bnow * fvec))

check_E_pi <- tibble(pi = seq(0, max_pi, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = list(
    tibble(
      E = E_vec_fxn(p = p, q = q, Xvec = Xvec, pi = pi, theta = theta, R = R, beta = beta, c = c),
      cty = names(fvec)
    )
  )) %>%
  unnest() %>%
  filter(!cty == "HS")
```

```{r}
ggplot(data = check_E_pi, mapping = aes(x = E, y = pi, color = cty)) +
  geom_line(size = 1) +
  geom_vline(xintercept = E, color = "red")
```

```{r}
check_E_pi %>%
  group_by(pi) %>%
  summarise(E = sum(E)) %>%
  ggplot(aes(x = E, y = pi)) +
  geom_line() +
  geom_vline(xintercept = E, color = "red") +
  geom_vline(xintercept = sum(max_Ei(p = p, q = q, Xvec = Xvec, theta = theta, R = R, beta = beta, c = c)[1:(n_patches - 1)]))
```

>>>>

---
title: "Run some checks"
author: "Villaseñor-Derbez J.C."
date: "Updated on `r Sys.Date()`"
output:
  html_document: 
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: false
---

# Model description

We model the PNA as a ten-patch discrete-time meta-population system, where Patch 1 is considering a spatial closure. Patches 2-9 are the other PNA countries, and Patch 10 represents the high seas and non-PNA countries. The stock of fish in each country is relatively stationary within a single fishing season, but redistributes across all patches annually. The price of fish is $p$, and catchability is given by $q$.

In the absence of a reserve, the revenue for vessels in patch $i$ is given by $pqE_iX_i$, where $E_i$ and $X_i$ are effort (vessel-days) and stock size in patch $i$ at the beginning of a period. The cost of fishing in patch $i$ is given by $cE_i^\beta$, where $\beta = 1.3$ matches commonly-used cost functions. Patch 1 considers a spatial closure by implementing a reserve as a fraction $R$ of the total patch ($R \in[0,1)$). Fish move within a patch based on $\theta$, where $\theta = 0$ implies no movement within the patch, and $\theta = 1$ implies that fish within the patch are well mixed during the fishing season. In this patch, revenues are given by $pqE_1X1(\theta + (1 - \theta)(1 - R))$. The parameterization of movement and reserve size imply that profit from fishing Patch 1 is given by:

$$
\begin{align}
\Pi_1(E_1,X_1,R) = pqE_1X_1(\theta + (1 - \theta)(1 - R))-cE_1^\beta
\end{align}
$$

And profits from fishing in Patch $j = \{2, 3, ..., 10\}$ are:

$$
\begin{align}
\Pi_j(E_j,X_j) = pqE_jX_j-cE_j^\beta
\end{align}
$$

The above equations imply that the marginal profit from the last unit of effort in a patch are given by:

$$
\begin{align}
\pi_1(E_1) = pqX_1(\theta + (1 - \theta)(1 - R)) - \beta cE_1^{\beta-1}
\label{eqn:marginal_profit_1}
\\
\pi_j(E_j) = pqX_j - \beta cE_j^{\beta-1}
\label{eqn:marginal_profit_j}
\end{align}
$$

In practice, the effort levels in each PNA patch are given (so $E_{1},\ E_{2},...,E_{9}$ are given) and the effort level on the high seas ($E_{10}$) is a result of open access dynamics. Therefore, we assume that effort continues to enter Patch 10 until the profit from the last unit of effort is exactly zero, indicating that $E_{10}$ is the value for which $\pi_{10}(E_{10})  = 0$. Setting Equation \ref{eqn:marginal_profit_j} equal to zero and solving for $E_j$ gives:

$$
\begin{align}
E_{10} = \left(\frac{pqX_{10}}{\beta c}\right)^{\frac{1}{(\beta - 1)}}
\label{eqn:effort_hs}
\end{align}
$$

The patch-level harvest is then determined by effort and stock size:

$$
\begin{align}
H_1 = qE_1X_1 (\theta + (1 - \theta)(1 - R))
\label{eqn:harvest_1}
\\
H_j = qE_jX_j
\label{eqn:harvest_j}
\end{align}
$$

Therefore, escapement in patch $i$ is the difference between initial stock size and harvests given by $e_{it} = X{it} - H_{it}$ and total escapement is $e_t=\sum_{i=1}^{10}e_{it}$. The entire stock then grows according to:

$$
\begin{align}
X_{t+1} = e_t + \frac{(\phi + 1)}{\phi}ge_t\left(1 - \left(\frac{e_t}{K}\right) ^{\phi}\right)
\label{eqn:grow}
\end{align}
$$

After the stock grows, a constant and patch-specific fraction $f_i$ of the total stock redistributes to patch $i$, so:

$$
\begin{align}
X_{it+1} = f_iX_{t+1}
\label{eqn:disperse}
\end{align}
$$

The vessel-day price that a country charges is given by $\pi_i$ from Equations \ref{eqn:marginal_profit_1} and \ref{eqn:marginal_profit_j}. Therefore, patch-level license revenues are given by:

$$
\begin{align}
\omega = \pi_iE_i
\label{eqn:license_revenue}
\end{align}
$$

Equations \ref{eqn:harvest_1} shows that low values of $\theta$ and $R > 0$ would increase escapement, which would lead to an increase in stock size (Equation \ref{eqn:grow}) and a benefit to all the other patches. But this would also cause the stock in the high seas ($X_{10}$) to increase, leading to an increased effort being allocated to the high seas (Equation \ref{eqn:effort_hs}) and a loss of these potential rents. Thus, the spillover benefits of increasing $R$ are never completely captured.

The vessel-day price for parcels that do not implement a reserve ($j = c(2, 10)$) is given by:

$$
\pi(E_j) = pqX_j - \beta c E_j ^{\beta - 1}
$$

We can solve for $E_j$:

$$
\begin{split}
\pi_j + \beta c E_j ^{\beta - 1} &= pqX_j \\
\beta c E_j ^{\beta - 1} &= pqX_j - \pi_j \\
E_j ^{\beta - 1} &= \frac{pqX_j - \pi_j}{\beta c } \\
E_j &= \left(\frac{pqX_j - \pi_j}{\beta c }\right) ^ { \frac{1}{\beta - 1}}
\end{split}
$$

For the parcel that implements the reserve ($i = 1$) the same equations are given by:

$$
\pi_1(E_1) = pqX_1 (\theta + (1 - \theta)(1 - R))- \beta c E_1 ^{\beta - 1}
$$

For simplicity, we can write $\Omega = (\theta + (1 - \theta)(1 - R))$, and the baove becomes

$$
\pi_1(E_1) = pqX_1\Omega - \beta c E_1 ^{\beta - 1}
$$

We can solve for $E_i$ as in the previous casse:

$$
\begin{split}
\pi_1 + \beta c E_1 ^{\beta - 1} &= pqX_1\Omega \\
\beta c E_1 ^{\beta - 1} &= pqX_1\Omega - \pi_1 \\
E_1 ^{\beta - 1} &= \frac{pqX_1\Omega - \pi_1}{\beta c} \\
E_1 &= \left(\frac{pqX_1\Omega - \pi_1}{\beta c }\right) ^ {\frac{1}{\beta - 1}}
\end{split}
$$

The total vessel-days for cells that belong to PNA countries (*i.e.* i = (1, 9)) are constrained at 45,000, which we term $\bar{E}$. Therefore, total effort can be expressed as:

$$
\bar{E} = \left(\frac{pqX_1(\theta + (1 - \theta)(1 - R)) - \pi_1}{\beta c }\right) ^ {\frac{1}{\beta - 1}} + 
\sum_{j = 2}^9\left(\frac{pqX_j - \pi_j}{\beta c }\right) ^ { \frac{1}{\beta - 1} }
$$

This parameterization shows that the cell-level effort is determined by the combination of stock size ($X_j$) and the price $\pi$. Trading between country will continue until all $\pi_i$ reach an equilibrium price of $\bar{\pi}$.

# Setup

@@

```{r}
# Load libraries
library(startR)
library(here)
library(cowplot)
library(magrittr)
library(tidyverse)

# Load functions
source(here("scripts", "model", "E_vec_fxn.R"))
source(here("scripts", "model", "pi_of_E.R"))
source(here("scripts", "model", "max_Ei.R"))

# Some housekeeping
update_geom_defaults("point",
                     list(
                       fill = "steelblue",
                       color = "black",
                       shape = 21,
                       size = 3
                     ))

update_geom_defaults("line", list(color = "black",
                                  size = 1))

theme_set(startR::ggtheme_plot(font_size = 8))
```

@@

Let's take a look at the relationship between Effort and Catches:

```{r}
left_join(act, catches, by = c("year", "eez_iso3")) %>%
  drop_na() %>% 
  ggplot(mapping = aes(x = days, y = catches, fill = year)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point() +
  labs(x = "Effort (days)", y = "Catches (tonnes)") +
  ggtitle("Relationship between effort and catches") +
  scale_fill_viridis_c()
```

```{r}
fvec_cpue <- c(cpue$cpue_prop, prop_hs)
fvec_cpue <- c(0.12, 0.08, 0.07, 0.13, 0.07, 0.12, 0.09, 0.12, 0.1, 0.1)
names(fvec_cpue) <- c(levels(cpue$eez_iso3), "HS")

@@

## Is there a value of $\bar{E}$ for the given possible ranges of Biomass?

The following plot shows the attainable levels of effort given price, catchability and biomass for cases where pi = 0. In other words, this just shows if harvesting at all would be profitable even in the absence of a vessel-day price.

The graph shows that for biomass levels between 1 ton and current Byomass (`r Bnow`), many values of $E$ are above $\bar{E}$ (horizontal dashed line)

```{r}
tibble(X = seq(1, Bnow, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = sum(max_Ei(p = p, q = q, Xvec = X, theta = theta, R = R, beta = beta, c = c))) %>%
  ggplot(aes(x = X, y = E)) +
  geom_line() +
  geom_vline(xintercept = Bmsy) +
  geom_hline(yintercept = E, linetype = "dashed") +
  labs(x = "Biomass", y = "Effort (days)")
```

For 95% of the values of biomass, there are attainable levels of effort that would produce positive catches.

```{r}
tibble(X = seq(1, Bnow, length.out = 100)) %>%
  rowwise() %>%
  mutate(E_calc = sum(max_Ei(p = p, q = q, Xvec = X, theta = theta, R = R, beta = beta, c = c)),
         Eabove = E_calc > E) %$%
  sum(Eabove)
```

## What's the aggregate demand for a given value of pi with Bmsy?

First we need to establish the maximum chargable price to plot all ranges from 0 to max_pi.

```{r}
max_pi <- max((p * q * Bnow * fvec))
```

For given values, the maximum chargable price for VDS is `r max_pi`. Let's create a dataframe with this data. This data.frame essentially contains the demand curves for each country, based on the biomass in each patch. Biomass comes from the cpue vector we derived.

```{r}
check_E_pi <- tibble(pi = seq(0, max_pi, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = list(
    tibble(
      E = E_vec_fxn(p = p, q = q, Xvec = 0.9 * Xvec, pi = pi, theta = theta, R = R, beta = beta, c = c),
      cty = names(fvec),
      cty2 = as.character(1:length(fvec))
    )
  )) %>%
  unnest() %>%
  filter(!cty == "HS")
```

The following graph shows all the demand curves for each country, given a value of E. The vertical dashed line shows the $\bar{E}$. Its intersections with each demand curve tell us how much vessels would be willing to pay to fish in each country.

```{r}
(demands <- ggplot(data = check_E_pi, mapping = aes(x = E / 1000, y = pi, color = cty2)) +
   geom_line(size = 1) +
   geom_vline(xintercept = E / 1000, color = "red", linetype = "dashed") +
   labs(x = "E (thousand vessel-days)", y = quo(pi)) +
   guides(color = guide_legend("Patch", ncol = 2)) +
   theme(legend.position = c(1, 1),
         legend.justification = c(1, 1)) +
   scale_color_brewer(palette = "Set1")
   )
```

We can aggregate all these demand curves horizontally and obtain the total demand curve. This is essentially the curve along wich we want to find the optimal pi.

```{r}
(demands_agg <- check_E_pi %>%
   group_by(pi) %>%
   summarise(E = sum(E)) %>%
   ggplot(aes(x = E / 1000, y = pi)) +
   geom_line() +
   geom_vline(xintercept = E / 1000, color = "red", linetype = "dashed") +
   labs(x = "E (thousand vessel-days)", y = quo(pi))
 )
```

```{r, fig.width = 7, fig.height = 3.5}
(demand_curves <- plot_grid(demands, demands_agg, ncol = 2))
```

```{r}
ggsave(plot = demand_curves,
       filename = here("docs", "slides", "img", "demand_curves.png"),
       width = 4,
       height = 2)
```

# Reproducibility chunk

```{r}
sessionInfo()
```
